Bootstrap learning via modular concept discovery

### 2013-02-01 drafts/ijcai2013/ijcai13.tex
Changes.

\documentclass{article}

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{float}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{alltt}
\usepackage{dashrule}
\usepackage{array}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}

@@

In this work, we try to see how far this idea of reusable function
definition can get us. Our question is, Can an AI system discover the
structure latent in the solutions to multiple related problems and
thereby bootstrap effective exploration of the search space? Can
discovering modular program components transform a problem from one in
which search is intractable into a problem in which it becomes
increasingly feasible with experience?

Concretely, our algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies.

@@

\section{Related Work}
The idea of discovering and using reusable subcomponents is an old one
in AI. Rendell's 1985 paper~\cite{DBLP:conf/ijcai/Rendell85},
``Substantial Constructive Induction Using Layered Information
Compression: Tractable Feature Formation in Search,'' -- which could
have been the name of this paper -- did not use programs as its
solution space representation. Within the Inductive Logic Programming
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. We believe that applying
constructive induction to a compositional functional programming
paradigm will provide even greater benefits.

@@

\section{Problem statement and terminology}
Our goal is to solve the following minimal feedback multi-task program
induction problem: suppose we are given set of \emph{tasks}
$T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
in our language.  We say that $ p \in \mathcal{L}$ \emph{solves} $t_i$
if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in $T$ as
we can.

We will use the following terminology. A \emph{frontier} will be the
finite set of programs that the algorithm is considering at any one
time. We will say that a task is \emph{hit} by the frontier if there
is a program in the frontier that solves it.

@@

In order to solve our stated problem, we need to choose a solution
representation language and a compression criterion, and we need to
describe the frontier enumeration and compression procedures.

@@

\subsection{A stochastic grammar over programs.}
\label{sec:stochgrammar}
We would like a way to describe how likely we believe a program to be
given the probabilities we ascribe to its components. Recall that
$\mathcal{L}$ is the set of all programs.  A stochastic grammar over
programs represents a distribution over expressions in
$\mathcal{L}$. We associate with each primitive combinator $c_i$ in
$\mathcal{L}$ a probability $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1
$.

The probability of using a primitive combinator somewhere in a program
should depend on which other combinators could have been used in its
place. This, in turn, depends on the type at that particular location
in the program tree; we call this type the \emph{requesting} type. For
example, suppose that we have the primitive $(+1)$ of type $Int
\rightarrow Int$ and the primitive $NOT$ of type $Bool \rightarrow
Bool$. If the requesting type is $\sigma \rightarrow Int$, then we can
only use $(+1)$, but if the requesting type is $\sigma \rightarrow
\sigma$ then we could use either $(+1)$ or $NOT$. Let $C = c_1, \dots,
c_n$ be the set of primitive combinators in our grammar; we define
$C_t$ for any type $t$ to be the set of combinators in $C$ that are
consistent with $t$, that is, whose types \emph{unify} with t (we use
unification according to the standard type theory
presentation~\ref{Pierce_2002}).

We define the probability of an expression $e$ from a grammar with
primitive $C$ to be $p(e)~=~\prod_{n=1}^N p(c_n | t(c_n))$ where
$p(c_n | t(c_n))$ is the probability of using primitive $c_n$ when the
requesting type was $t(c_n)$. This definition means that the
probability of an expression only depends on the probabilities
assigned to the library primitives. We define these conditional
probabilities as
\[
p(c_i | t(c_n)) =    \frac{1}{\epsilon}\frac{p_i}{\sum_{C_{t(c_n)}} p_j}.
\]
That is, $p(c_i | t(c_n)$ is the probability of sampling $c_i$ from
the multinomial distribution defined by the $p_i$'s, conditioned on
the requesting type, with the factor of $\frac{1}{\epsilon}$ added as
a simple way to ensure that the sum over the probabilities of all
expressions converges to a finite value (so that the distribution over
expressions is proper). This is true whenever $\epsilon \geq
4$\footnote{The probability mass $M_d$ of all expressions of depth
  less than or equal to $d$ can be written with the recurrence
  relation $M_d \leq M_{d-1}^2 + M_1$. This has an asymptotic fixed
  point, as $d \rightarrow \infty$ if $x = x^2 + M_1$ has a solution,
  which is true if $M_1 \leq 1/4$.}.

To estimate the values $p_i$ from a set of expressions, we simply divide
the number of times primitive combinator $c_i$ was used by the number
of times it could have been used given the requesting types in the
expression trees. Note that this is not a maximum likelihood estimator
for this distribution; finding that would require solving a rather
complex optimization problem because of each choice is conditioned on
the requesting type. We choose our estimator merely for convenience,
and find that our empirical results justify this simplicity. It is a
matter for future research to determine to what extent a more accurate
estimator would improve performance.

\subsection{Best-first enumeration of most probable programs.}
In order to explore the frontier of most promising programs, we need a
procedure that enumerates the $N$ most probable programs. There has
been recent interest in this problem, most of it focused on
enumerating the shortest program satisfying some
criterion~\cite{DBLP:conf/sfp/Katayama05}
\cite{DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
problem as a best-first exploration of an AND/OR tree
(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73},
which we denote $\mathcal{G}$. In words, the enumeration procedure is
best described by the following recursion: every program is either a
primitive combinator or it a left child program applied to a right
child program.

This translates to the following AND/OR tree. Suppose we want a
program of type $t$. The root of $\mathcal{G}$ is an OR node of type
$t$. Its children are elements in $C_t$ (defined in the previous
section) and one AND node of type $t$ with two children. Each child of
an AND node has the same structure as the root node, with modified
types: type of its left child is $\tau \rightarrow t$ (where $\tau$ is
a fresh type variable). Since the type of the right child is
constrained by the subtree rooted at the left child, we always expand
the left child first. Once we have a complete left child program, we
can use its target type as the type of the right child.

@@

Having enumerated the frontier, we want to assign solutions
to hit tasks such that this set is maximally compressible. This will
promote the usage of reusable modular subtrees.

@@

The top row consists of one type of representation of an integer
squared, namely, the integer times itself. The bottom row contains
another type of representation in which the function $f(x)=x^2$,
represented in combinatory logic as $S * I$, is applied to the
integer. The top and bottom row of each expression evaluates to the
same integer. If we consider only the first column, the top tree has 4
unique subtrees and the bottom has 7. But each column adds 3 unique
subtrees to the top row and only 2 to the bottom. So by the fifth
column the top row has more unique subtrees than the bottom row. The
increased compressibility of the bottom row is due to the fact that
the squaring operation has been factored out as an abstraction. This
is an example, then, of how this compressibility metric favors
abstraction when multiple expressions are being considered at once.

@@

Given a set of solution programs, we want to re-estimate our
stochastic grammar. Our inspiration for doing this is based on
the Neville-Manning algorithm for grammar-based
compression~\cite{nevill1997identifying}. The idea of that compression
algorithm is to compress any string by creating a grammar for that
string such that a) no two symbols appear more than once in the
grammar and b) every rule is used more than once. From a set of trees
we generate a grammar according to these criteria.

@@

In this section, we discuss the performance of our learning algorithm
on a symbolic regression problem. In this problem, each task is a set
of pairs $ (x, f(x))$ where $x \in (0, \dots, 9)$ and $f$ is a
polynomial. The goal is to discover a program for each task that
outputs the appropriate value $f(x)$ for each input x.

%% The choice of an initial grammar for the algorithm will, of course,
%% have a big effect on its performance, and so we would like to
%% characterize this effect. We can divide primitives into roughly two
%% sets; there are low-level primitives, those that operate on concrete
%% types like Bool or Double, and these tend to be domain specific. For
%% example, concrete primitives include $1, 2$, etc., and $f(x) = 3x$. The
%% other class of primitives are abstract primitives, which can be
%% characterized by having arguments whose types are primarily function
%% types, like $(t \rightarrow s), (t \rightarrow s) \rightarrow \text{
%%   Double },$ etc. Of course, we can construct operators that fall in
%% between.

%% The most abstract operators are those that Liang et. al refer to as
%% ``routers''~\cite{liang10programs}. These operators take the place of
%% $\lambda$ abstraction in the general lambda calculus: whereas
%% $\lambda$ abstraction allows us to bind arguments to variables nested
%% deep inside an expression, without them we need to have operators that
%% ``route'' arguments to the appropriate functions in a nested
%% expression.

%% \begin{definition}[Router~\cite{liang10programs}] A \emph{router} is a combinator
%% represented by a finite sequence of elements from $\{B, C, S\}$. The
%% behavior of router $r$ is given by
%% \[
%% (r \, f \, g \, x_1 \dots x_{|r|}) = (f \, x_{i_1} \dots x_{i_n}) (g
%% \, x_{j_1} \dots x_{j_n}),
%% \]
%% where $i_1 < \dots < i_n$ are indices $i$ such that $r_i \in \{C, S\}$
%% and where $j_1 < \dots < j_n$ are indices $j$ such that $r_j \in \{B,
%% S\}$. Let $\mathcal{R}_k$ be the set of all routers of length equal to
%% $k$. Let $\mathcal{R}_{\leq k}$ be the set of all routers of length
%% less than or equal to $k$.
%% \end{definition}

%% To analyze the behavior of our algorithm, we will vary the parameters
%% of the problem along two axes while keeping the learning tasks the
%% same. Along the first axis, we vary the initial library to which our
%% learner has access with respect to the its abstractness (i.e., number
%% of routers) and its concreteness (i.e. number of concrete
%% primitives). Along the second axis we vary the size of the search
%% frontier.

@@

\begin{table}
\tiny
\begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
\hline \\
  function  & CL expression & circuit  ~ \\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

%\begin{top}
\begin{circuitikz}[transform shape] 
\draw
(1, 0) node[nand port, scale=.5] (nand1) {}
(2,.3) node[nand port, scale=.5] (nand2) {}
(0, 0) node[anchor=east]  (x) {x}
(x) -- (.3, 0) -| (nand1.in 1)
(x) -- (.3, 0) -|  (nand1.in 2)
(x) |- (nand2.in 1)
(nand1.out) |- (nand2.in 2)
;
\end{circuitikz}
%\end{top}
\\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~

%% \begin{circuitikz}[transform shape]

%% ;
%% \end{circuitikz}

\\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~

\begin{circuitikz}[transform shape]
\draw
(1.2,0) node[nand port, scale=.5] (nand1) {}
(0, 0) node (x) {x}
(x) -| (nand1.in 1)
(x) -| (nand1.in 2)
;
\end{circuitikz}

\\ \hline
¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

\begin{circuitikz}
\draw
(1.0,0) node[nand port, scale=.5] (nand1) {}
(2.0,0) node[nand port, scale=.5] (nand2) {}
% (0, 0) node (x) {x}
% (0, 0) node (y) {y}
node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
(x) |- (nand1.in 1)
(y) |- (nand1.in 2)
(nand1.out) |- (nand2.in 1)
(nand1.out) |- (nand2.in 2)
;
\end{circuitikz}
 \\  \hline
¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~

\begin{circuitikz}
\draw
(0,0) node (x) {x}
(0,.7) node (y) {y}

node[nand port, scale=.5, right=1cm of x] (nand1) {}
node[nand port, scale=.5, right=1cm of y] (nand2) {}
node[nand port, scale=.5, below right=.3cm and 1cm of nand2] (nand3) {}
(x) -| (nand1.in 1)
(x) -| (nand1.in 2)
(y) -| (nand2.in 1)
(y) -| (nand2.in 2)
(nand1.out) |- (nand3.in 1)
(nand2.out) |- (nand3.in 2)
;
\end{circuitikz}
 \\  \hline
¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

\begin{circuitikz}
\draw
node[nand port, scale=.5] (nand1) {}
node[nand port, scale=.5, above right=.3cm and 1cm of nand1] (nand2) {} 
node[nand port, scale=.5, below right=.3cm and 1cm of nand1] (nand3) {} 
node[nand port, scale=.5, right=2cm of nand1] (nand4) {} 
node[left=1.2cm of nand2.in 1] (x) {x}
node[left=1.2cm of nand3.in 2] (y) {y}
node[right=.2cm of nand4] (out) {}
(x) -| (nand1.in 1)
(x) -| (nand2.in 1)
(y) -| (nand1.in 2)
(y) -| (nand3.in 2)
(nand1.out) -| (nand2.in 2)
(nand1.out) -| (nand3.in 1)
(nand2.out) |- (nand4.in 1)
(nand3.out) |- (nand4.in 2)
(nand4.out) -- (out)
;
\end{circuitikz}

 \\  \hline
¬†¬†¬†¬†¬†¬†¬†¬†F¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~
%% \begin{circuitikz}
%% \draw
%% (0, 0) node[left=.25) (x) {x}

%% ;
%% \end{circuitikz}

\\ 
\hline

\end{tabular}
\caption{Expressions in the induced library on sets of tasks
  corresponding to boolean circuits synthesized from AND, OR, and NOT
  gates, with boolean circuits of 1, 2, or 3 inputs.}
\label{table:boolexpr}

\end{table}

>>>>

x\documentclass{article}

\usepackage{ijcai13}
\usepackage{times}
\usepackage{amsmath} 
\usepackage{latexsym} 
\usepackage{ dsfont }
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{tikz-qtree}
\usepackage{float}
\usepackage{circuitikz}
\usepackage{graphicx}
\usepackage[font=small]{subcaption}
\usepackage[font=small]{caption}
\usepackage{alltt}
\usepackage{dashrule}
\usepackage{array}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usetikzlibrary{arrows,automata, positioning, patterns,backgrounds}
\usetikzlibrary{arrows,shapes.gates.logic.US,shapes.gates.logic.IEC,calc}

@@

In this work, we try to see how far this idea of reusable function
definition can get us. We ask the following question: can an AI system
discover the structure latent in the solutions to multiple related
problems and thereby bootstrap effective exploration of the search
space? Can discovering modular program components transform a problem
from one in which search is intractable into a problem in which it
becomes increasingly feasible with experience?

Concretely, our algorithm works as follows. It begins with a small
library of program primitives and a distribution over programs
synthesized from this library. In each iteration, it generates
programs in order of highest probability according to the library
learned in the previous iteration. From the discovered solutions, it
finds new reusable functions that compress these solutions, and it
re-estimates the distribution over programs by weighting these new
functions according to their frequencies. In using compression as a
guide to the choice representation, we instantiate the idea that good
representations are those which minimize the description length of
typical elements in the domain.

@@

\section{Related Work}
The idea of discovering and using reusable subcomponents is an old one
in AI. Rendell's 1985 paper~\cite{DBLP:conf/ijcai/Rendell85},
``Substantial Constructive Induction Using Layered Information
Compression: Tractable Feature Formation in Search,'' contained many
of these ideas. It did not, however, use programs as its solution
space representation. Within the Inductive Logic Programming
literature, researchers have explored ``Constructive Induction,''
generating new and reusable logic programming
terms~\cite{DBLP:conf/ijcai/Muggleton87}. We believe that applying
constructive induction to a compositional functional programming
paradigm will provide even greater benefits.

@@

@@

Our goal is to solve the following minimal feedback multi-task program
induction problem: suppose we are given set of \textbf{tasks}
$T=\{t_k\}_{k=1}^K$ where each task is a function $t_i : \mathcal{L}
\rightarrow \{0, 1\}$ and where $\mathcal{L}$ is the set of programs
in our language.  We say that $ p \in \mathcal{L}$ \textbf{solves}
$t_i$ if $ t_i(p) = 1$. Our goal is to solve as many of the tasks in
$T$ as we can.

We propose the \textbf{E-C algorithm}, an iterative procedure to solve this
problem. It will be useful in presenting the algorithm to define the
following terminology: a \textbf{frontier} is the finite set of
programs that the algorithm considers in any one iteration. The
\textbf{frontier size} $N$ is the prespecified number of elements in
the frontier. We will say that a task is \textbf{hit} by the frontier
if there is a program in the frontier that solves it.

The E-C algorithm maintains a distribution $\mathcal{D}$ over
expressions in $\mathcal{L}$. At each iteration, it
\begin{enumerate}
\item \textbf{explores} the frontier -- enumerating the $N$ most probable expressions
  from the current distribution $\mathcal{D}$ -- and
\item \textbf{compresses} the hit tasks in the frontier to estimate a new
  distribution.
\end{enumerate}

What remains is to specify a representation language $\mathcal{L}$, a
form for the distribution $\mathcal{D}$, a compression criterion for a
set of expressions in $\mathcal{L}$. We also need to describe
procedures for the enumeration of the frontier, compression of the
expressions, and re-estimations of $\mathcal{D}$.

@@

\subsection{A stochastic grammar over programs}
\label{sec:stochgrammar}

We specify a distribution $\mathcal{D}$ over expressions in
$\mathcal{L}$ as a stochastic
grammar~\cite{feldman1969grammatical}. Many stochastic grammars have
the desirable property that the probability of an expression is the
product of the probabilities of its subcomponents. The distribution we
specify will be a particularly simple verson of this, in which the
probability of an expression is the product of the probability of its
leaves.

$\mathcal{D}$ will associate with each primitive combinator $c_i$ a
prior probabilities $p_i$, $0~\leq~p_i~\leq~1, \sum_i~p_i~=~1 $. The
probability of a leaf withr primitive combinator $c_i$ will depend on
which other primitive combinators could have been chosen in its
place. This set will depend on the type that was requested by the
parent of this leaf when the leaf's primitive was chosen; we call this
type the \emph{requesting} type of the leaf. For example, suppose that
we have the primitive $(+1)$ of type $Int \rightarrow Int$ and the
primitive $NOT$ of type $Bool \rightarrow Bool$. If the requesting
type is $\sigma \rightarrow Int$, then we can only use $(+1)$, but if
the requesting type is $\sigma \rightarrow \sigma$ then we could use
either $(+1)$ or $NOT$. Let $C = c_1, \dots, c_n$ be the set of
primitive combinators in our grammar; we define $C_t$ for any type
$\tau$ to be the set of combinators in $C$ that are consistent with
$\tau$, that is, whose types \emph{unify} with $\tau$ (we use
unification according to the standard type theory
presentation~\ref{Pierce_2002}).

Given a distribution $\mathcal{D} = (p_1, \dots, p_n)$, we define the
probability of an expression $e \in \mathcal{L}$ to be
$p(e)~=~\prod_{n=1}^N p(c_n | \tau(c_n))$ where $p(c_n | \tau(c_n))$ is the
probability of using primitive $c_n$ when the requesting type is
$\tau(c_n)$. In turn, we define the conditional probabilities to be
\[
p(c_i | \tau(c_n)) \propto \frac{p_i}{\sum_{C_{\tau(c_n)}} p_j}.
\]
That is, $p(c_i | \tau(c_n)$ is proportional to the probability of
sampling $c_i$ from the multinomial distribution defined by the
$p_i$'s, conditioned on the requesting type, with the constant of
proportionality chosen to ensure that the sum over the probabilities
of all expressions converges to a finite value (so that the
distribution over expressions is proper). For binary trees, this is
true whenever the constant of proportionality is less than
$\frac{1}{4}$\footnote{The probability mass $M_d$ of all expressions
  of depth less than or equal to $d$ can be written with the
  recurrence relation $M_d \leq M_{d-1}^2 + M_1$. This has an
  asymptotic fixed point, as $d \rightarrow \infty$ if $x = x^2 + M_1$
  has a solution, which is true if $M_1 \leq 1/4$.}.

Now, suppose that we have a set of expressions $\mathcal{E}$, and we
wish to estimate the maximum likelihood values of the $p_i$'s. If the
choice of primitive at each leaf were not conditioned on the
requesting type, we could just count all occurences of each primitive
in our expression set, and this would be proportional to the maximum
likelihood estimate. This is ML estimator for a multinomial
distribution. However, the fact that we condition on the requesting
type makes this a considerably more difficult optimization task. One
of the simplest approximations we can make is to calculate the
frequency with which each primitive combinator appears in
$\mathcal{E}$ when the requesting type is such that it could have
been chosen. This is a particularly straightforward calculuation, and
we use it for its convenience, finding that our empirical results
justify its use. Future inquiry will be needed to determine to what
extent a more accurate estimator can improve these results.

\subsection{Best-first enumeration of programs}
In order to explore the frontier of most promising programs, we need a
procedure that enumerates the $N$ most probable programs. There has
been recent interest in this problem, most of it focused on
enumerating the shortest program satisfying some
criterion~\cite{DBLP:conf/sfp/Katayama05}
\cite{DBLP:conf/aaip/YakushevJ09}. Our approach is to formulate the
problem as a best-first exploration of an AND/OR tree
(Figure~\ref{fig:andor})~\cite{nilsson1982principles}~\cite{DBLP:journals/cacm/Hall73},
which we denote $\mathcal{G}$. In words, the enumeration procedure is
best described by the following recursion: every program is either a
primitive combinator or it is a left child program applied to a right
child program.

This can framed as the following AND/OR tree. Suppose we want a
program of type $t$. The root of $\mathcal{G}$ is an OR node of type
$t$. Its children are elements in $C_t$ (defined in the previous
section) and one AND node of type $t$ with two children. Each child of
an AND node has the same structure as the root node, with modified
types: type of its left child is $\tau \rightarrow t$ (where $\tau$ is
a fresh type variable). Since the type of the right child is
constrained by the subtree rooted at the left child, we always expand
the left child first. Once we have a complete left child program, we
can use its target type as the type of the right child.

@@

Having enumerated the frontier, we want to assign solutions
to hit tasks such that this set is maximally compressible. This will
promote reusable modular subtrees.

@@

The top row consists of one type of representation of an integer
squared, namely, the integer times itself. The bottom row contains
another type of representation in which the function $f(x)=x^2$,
represented in combinatory logic as $S * I$, is applied to the
integer. The top and bottom row of each expression evaluates to the
same integer. If we consider only the first column, the top tree has 4
unique subtrees and the bottom has 7. But each column adds 3 unique
subtrees to the top row and only 2 to the bottom. So by the fifth
column the top row has more unique subtrees than the bottom row. The
increased compressibility of the bottom row is due to the fact that
the squaring operation has been factored out as an abstraction. This
is an example of how this compressibility metric favors
abstraction when multiple expressions are being considered at once.

@@

Given a set of solution programs, we want to re-estimate our
stochastic grammar. Our inspiration for doing this is the
Neville-Manning algorithm for grammar-based
compression~\cite{nevill1997identifying}. The idea of that compression
algorithm is to compress any string by creating a grammar for that
string such that a) no two symbols appear more than once in the
grammar and b) every rule is used more than once. From a set of trees
we generate a grammar according to these criteria.

@@

We first illustrate the performance of our learning algorithm on a
symbolic regression problem. In task in this problem, we want to find
an algebraic function, symbolically specified, that maps a set of
input values to output values. We choose this problem because it has a
long history in AI, particularly in the genetic programming
literature~\cite{DBLP:books/daglib/0070933}.

In the formulation we consider, each task $t$ corresponds to a
polynomial $f$, and an expression $e$ solves $t$ if ${ (e\;
  i)=f(i) }$ for all ${ i=0,\dots,9 }$.

@@

\include{./figures/tablealone}
%% \begin{table}
%% \tiny
%% \begin{tabular}[width=\linewidth]{|p{1cm}|p{2cm}|p{3.5cm}|}
%% \hline function  & CL expression & circuit  ~ \\ \hline\hline
%% ¬†¬†¬†¬†¬†¬†¬†¬†NOT¬†¬† & S NAND I¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~

%% \begin{circuitikz}[transform shape]
%% \draw
%% (1.2,0) node[nand port, scale=.3] (nand1) {}
%% (0, 0) node (x) {x}
%% (x) -| (nand1.in 1)
%% (x) -| (nand1.in 2)
%% ;
%% \end{circuitikz}

%% \\ \hline
%% ¬†¬†¬†¬†¬†¬†¬†¬†AND¬†¬† & ((C B NAND) (B (S NAND I)))¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

%% \begin{circuitikz}
%% \draw
%% (1.0,0) node[nand port, scale=.3] (nand1) {}
%% (2.0,0) node[nand port, scale=.3] (nand2) {}
%% % (0, 0) node (x) {x}
%% % (0, 0) node (y) {y}
%% node[above left=.2cm and .7cm  of nand1] (x) {x} %-- (nand1.in 1)
%% node[below left=.2cm and .7cm  of nand1] (y) {y} %-- (nand1.in 2)
%% (x) |- (nand1.in 1)
%% (y) |- (nand1.in 2)
%% (nand1.out) |- (nand2.in 1)
%% (nand1.out) |- (nand2.in 2)
%% ;
%% \end{circuitikz}
%%  \\  \hline
%% ¬†¬†¬†¬†¬†¬†¬†¬†OR¬†¬†¬† & ((B (C (B (S NAND) NAND))) (S NAND I) & ~

%% \begin{circuitikz}
%% \draw
%% (0,0) node (x) {x}
%% (0,.7) node (y) {y}

%% node[nand port, scale=.3, right=1cm of x] (nand1) {}
%% node[nand port, scale=.3, right=1cm of y] (nand2) {}
%% node[nand port, scale=.3, below right=.3cm and 1cm of nand2] (nand3) {}
%% (x) -| (nand1.in 1)
%% (x) -| (nand1.in 2)
%% (y) -| (nand2.in 1)
%% (y) -| (nand2.in 2)
%% (nand1.out) |- (nand3.in 1)
%% (nand2.out) |- (nand3.in 2)
%% ;
%% \end{circuitikz}
%%  \\  \hline\hline

%% ¬†¬†¬†¬†¬†¬†¬†¬†TRUE¬† & (S NAND) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

%% %\begin{top}
%% \begin{circuitikz}[transform shape] 
%% \draw
%% (1, 0) node[nand port, scale=.3] (nand1) {}
%% (2,.3) node[nand port, scale=.3] (nand2) {}
%% (0, 0) node[anchor=east]  (x) {x}
%% (x) -- (.3, 0) -| (nand1.in 1)
%% (x) -- (.3, 0) -|  (nand1.in 2)
%% (x) |- (nand2.in 1)
%% (nand1.out) |- (nand2.in 2)
%% ;
%% \end{circuitikz}
%% %\end{top}
%% \\ \hline
%% ¬†¬†¬†¬†¬†¬†¬†¬†FALSE & (S B (S NAND)) (S NAND I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† &  ~

%% \begin{circuitikz}[transform shape] 
%% \draw
%% (1, 0) node[nand port, scale=.3] (nand1) {}
%% (2,.3) node[nand port, scale=.3] (nand2) {}
%% node[nand port, scale=.3, right=1cm of nand2] (nand3) {} 
%% (0, 0) node[anchor=east]  (x) {x}
%% (x) -- (.3, 0) -| (nand1.in 1)
%% (x) -- (.3, 0) -|  (nand1.in 2)
%% (x) |- (nand2.in 1)
%% (nand1.out) |- (nand2.in 2)
%% (nand2.out) -| (nand3.in 1)
%% (nand2.out) -| (nand3.in 2)
%% ;
%% \end{circuitikz}
%% \\ \hline
%% ¬†¬†¬†¬†¬†¬†¬†¬†XOR¬†¬† & ((S (B C (((B (C B NAND)) S) (((B (C B NAND)) S) (B C ((B (B NAND)) NAND)))))) I)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

%% \begin{circuitikz}
%% \draw
%% node[nand port, scale=.3] (nand1) {}
%% node[nand port, scale=.3, above right=.3cm and 1cm of nand1] (nand2) {} 
%% node[nand port, scale=.3, below right=.3cm and 1cm of nand1] (nand3) {} 
%% node[nand port, scale=.3, right=2cm of nand1] (nand4) {} 
%% node[left=1.2cm of nand2.in 1] (x) {x}
%% node[left=1.2cm of nand3.in 2] (y) {y}
%% node[right=.2cm of nand4] (out) {}
%% (x) -| (nand1.in 1)
%% (x) -| (nand2.in 1)
%% (y) -| (nand1.in 2)
%% (y) -| (nand3.in 2)
%% (nand1.out) -| (nand2.in 2)
%% (nand1.out) -| (nand3.in 1)
%% (nand2.out) |- (nand4.in 1)
%% (nand3.out) |- (nand4.in 2)
%% (nand4.out) -- (out)
%% ;
%% \end{circuitikz}

%%  \\  \hline\hline
%% ¬†¬†¬†¬†¬†¬†¬†¬†$E_1$¬†¬†¬†  & S B (S NAND)¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬†¬† & ~

%% \begin{circuitikz}
%% \draw
%% node (x) {x}
%% node[nand port, scale=.3, right=1.5cm of x] (nand) {} 
%% node[draw=blue!75, fill=blue!20, thick, 
%%   above right=.3cm and .2cm of x, anchor=west] (f1) {$g(x)$} 
%% node[draw=blue!75, fill=blue!20, thick, right=.5cm of nand] (f2) {$g(x)$}
%% node[right=.1cm of f2] (out) {}
%% (x) -| (f1.west)
%% (x) -| (nand.in 2)
%% (f1) -| (nand.in 1)
%% (nand.out) -- (f2)
%% (f2) -- (out);
%% \end{circuitikz}

%% \\ \hline
%% $E_2$¬†¬†¬†  & B (C B NAND) S & ~

%% \begin{circuitikz}
%% \draw
%% node (x) {x} 
%% node[nand port, scale=.3, right=1.5cm of x] (nand) {} 
%% node[draw=blue!75, fill=blue!20, thick, 
%%   above right=.2cm and .3cm of x, anchor=west] (g) {$g(x)$} 
%% node[draw=blue!75, fill=blue!20, thick, 
%%   below right=.2cm and .3cm of x, anchor=west] (f) {$f(x)$} 
%% (x) -- [right=.25] (f.west)
%% (x) |- (g.west)
%% (f.east) -| (nand.in 1)
%% (g.east) -| (nand.in 2)
%% (nand.out) -- (out)
%% (x)
%% ;
%% \end{circuitikz}

%% \\ \hline

%% \end{tabular}
%% \caption{Expressions in the induced library on sets of tasks
%%   corresponding to boolean circuits synthesized from AND, OR, and NOT
%%   gates, with boolean circuits of 1, 2, or 3 inputs.}
%% \label{table:boolexpr}

%% \end{table}
